Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-08 be.tex
conclusions 16

\begin{figure}
\includegraphics[width=\columnwidth]{zetaCrit.pdf}
\caption{\label{fig:zetaCrit} {\emph Top panel:} Critical value
  $\zeta_c$ of the wind heating parameter $\zeta \equiv
  \sqrt{1+(v_w/\sigma_0)^2}$ above which the Bernoulli parameter of
  the flow $\mathcal{B}$ is positive near the break radius for some
  value of $\rs$, calculated for core ($\Gamma=0.1$, dot-dashed line)
  and cusp ($\Gamma=0.8$, solid line) galaxies using
  equation~\eqref{eq:enthAnal}.. {\emph Bottom Panel:} Minimum value
  of the stagnation radius for which $\mathcal{B} > 0$ at the break
  radius as a function of $\zeta/\zeta_c < 1$. }
\end{figure}

>>>>

\begin{figure}
\includegraphics[width=\columnwidth]{zetaCrit.pdf}
\caption{\label{fig:zetaCrit} {\emph Top panel:} Critical value
  $\zeta_c$ of the wind heating parameter $\zeta \equiv
  \sqrt{1+(v_w/\sigma_0)^2}$ above which the Bernoulli parameter of
  the flow $\mathcal{B}$ is positive near the break radius for some
  value of $\rs$, calculated for core ($\Gamma=0.1$, dot-dashed line)
  and cusp ($\Gamma=0.8$, solid line) galaxies using
  equation~\eqref{eq:enthAnal}. {\emph Bottom Panel:} Minimum value
  of the stagnation radius for which $\mathcal{B} > 0$ at the break
  radius as a function of $\zeta/\zeta_c < 1$. }
\end{figure}

### 2015-04-08 ms.tex
conclusions 16

We have calculated steady-state models for the hot gaseous circumnuclear media of quiescent galaxies, under the assumption that gas is supplied exclusively by stellar wind mass loss and heated by shocked stellar winds, supernovae and black hole feedback.  We compute solutions for a range of different black hole masses, heating rates, and observationally-motivated stellar density profiles.  Our numerical results (Table \ref{table:models}) are used to verify and calibrate analytic relationships (Appendix \ref{app:rs}, \ref{app:be}).  The latter are in turn used to explore systematically how the SMBH accretion rate varies with black hole mass and the galaxy's SFH.  Our results for $\dot{M}(\Mbh)$ are compared with observed trends of the nuclear X-ray luminosities of quiescent SMBHs and low luminosity AGN.  They are also used to address key questions such as the thermal stability of the accretion flow.  Our conclusions are summarized as follows.

@@

\item Including the effects of heat transport by electron conduction results in at most order unity changes to the key properties of the flow (e.g. stagnation radius, mass accretion rate, and thermal stability) for causal values of the conduction saturation parameter $\phi < 0.1$ (eq.~[\ref{eq:conduction}]).  Angular momentum is also unlikely to invalidate our spherically symmetric models given the extrapolated rotation curves of galaxies to the nuclear scales of interest ($\S\ref{sec:rotation}$).

\item Radiative cooling has a more pronounced influence on the structure of the gas flow.  Cooling instabilities are likely to develop when the radiative cooling rate exceeds the gas heating rate near the stagnation radius.  This condition is approximately equivalent to the more criterion $t_{\rm cool} \gg t_{\rm ff}$ criterion advocated by \citet{McCourt+12} and \citet{Li&Bryan14a}.

@@

\item Stellar wind heating, supernovae, and AGN feedback depend explicitely on the SMBH mass as well as the SFH in the galactic nuclear region.  A young starburst of age $\lesssim 10$ Myr produces mass and energy injection dominated by young stellar winds ($v_w \sim 1000$ km s$^{-1}$), while for exclusively old stellar populations  mass input is dominated by AGB winds and heating is dominated by AGN feedback and supernovae.  Ongoing, continuous star formation presents a hybrid situation, with energy input usually dominated by young stars and mass input dominated by the old stars (Appendix \ref{app:windheat}).  Galactic nuclei with $\Mbh \lesssim 10^8\Msun$ heated according to their average SFHs (as derived from cosmological simulations) receive a stellar wind heating of $v_w \sim 700$ km s$^{-1}$, although this number can be somewhat lower if the star formation is intermittent (Fig.~\ref{fig:NickPlot2}).

\item Type Ia SNe only provide a continuous heating source exterior to the Ia radius $\rIa$ (eq.~[\ref{eq:rIa}]) where the time between subsequent SNe exceeds the dynamical timescale.  Heating from the average SFH results in $\rIa < \rs$, except for the most massive galaxies (Fig.~\ref{fig:vwSources}).  Nevertheless, even if other sources of heating are low enough for $\rs$ to approach $\rIa$, the SN Ia heating supplied when $\rs \gtrsim \rIa$ is sufficiently large that $\rs$ is unlikely to greatly exceed $\rIa$.  SNe Ia thus cap the SMBH accretion rate at the value $\dot{M}_{\rm Ia}$ (eq.~[\ref{eq:eddr_Ia}]), except possibly in high mass elliptical galaxies with large break radii, for which $\zeta(v_w^{\rm Ia}) \gtrsim \zeta_c$.

\item SMBH feedback, unlike stellar heating sources, depends on the SMBH accretion rate, which is itself a function of the heating rate.  Compton heating is generally unimportant compared to stellar wind heating when the latter corresponds to the average SFH; however, Compton heating can dominate in the case of an impulsive burst of star formation (\ref{fig:vwSourcesImp}).  In principle the Compton heating rate is sufficiently high to prevent thermal instabilities in steady-state, i.e. $\dot{M}_{C} \lesssim \dot{M}_{\rm TI}$ (Fig.~\ref{fig:bh_growth})  However, unlike local stellar feedback
mechanisms, SMBH feedback may not be capable of truly stabilizing the flow given its inability to respond instantaneously to local changes in the properties of the gas.  Kinetic feedback is more difficult to quantify, but could also be important in determining $\dot{M}(M_{\bullet}$) when stellar heating is weak ($\S\ref{sec:kinetic}$).

>>>>

We have calculated steady-state models for the hot gaseous circumnuclear media of quiescent galaxies, under the assumption that gas is supplied exclusively by stellar wind mass loss and heated by shocked stellar winds, supernovae and black hole feedback.  We compute solutions for a range of different black hole masses, heating rates, and observationally-motivated stellar density profiles.  Our numerical results (Table \ref{table:models}) are used to verify and calibrate analytic relationships (Appendix \ref{app:rs}, \ref{app:be}).  We use the latter to explore systematically how the SMBH accretion rate varies with black hole mass and the galaxy's SFH.  Our results for $\dot{M}(\Mbh)$ are compared with observed trends of the nuclear X-ray luminosities of quiescent SMBHs and low luminosity AGN.  They are also used to address key questions such as the thermal stability of the accretion flow.  Our conclusions are summarized as follows.

@@

\item Including the effects of heat transport by electron conduction results in at most order unity changes to the key properties of the flow (e.g. stagnation radius, mass accretion rate, and thermal stability) for causal values of the conduction saturation parameter $\phi < 0.1$ (eq.~[\ref{eq:conduction}]).  Angular momentum is also unlikely to invalidate our spherically symmetric model, given the inferred stellar rotation rates near the stagnation radius ($\S\ref{sec:rotation}$).

\item Radiative cooling has a more pronounced influence on the flow structure.  Cooling instabilities are likely to develop when the radiative cooling rate exceeds the gas heating rate near the stagnation radius.  This condition is approximately equivalent to the more criterion $t_{\rm cool} \gg t_{\rm ff}$ criterion advocated by \citet{McCourt+12} and \citet{Li&Bryan14a}.

@@

\item Stellar wind heating, supernovae, and AGN feedback depend explicitely on the SMBH mass as well as the SFH in the galactic nuclear region.  A young starburst of age $\lesssim 10$ Myr produces mass and energy injection dominated by young stellar winds ($v_w \sim 1000$ km s$^{-1}$), while for exclusively old stellar populations  mass input is dominated by AGB winds and heating is dominated by AGN feedback and supernovae.  Ongoing, continuous star formation presents a hybrid situation, with energy input usually dominated by young stars and mass input dominated by the old stars (Appendix \ref{app:windheat}).  Galactic nuclei with $\Mbh \lesssim 10^8\Msun$ heated according to their average SFHs (as derived from cosmological simulations) receive a stellar wind heating of $v_w \sim 700$ km s$^{-1}$.  This heating can be suppressed if the star formation is sufficiently intermittent (Fig.~\ref{fig:NickPlot2}).

\item Type Ia SNe only provide a continuous heating source exterior to the Ia radius $\rIa$ (eq.~[\ref{eq:rIa}]) where the time between subsequent SNe exceeds the dynamical timescale.  Non-Ia heating due to the average SFH results in $\rs <\rIa$, except for the most massive galaxies (Fig.~\ref{fig:vwSources}).  However, if $\rs$ does approach or exceed $\rIa$, then the SN Ia heating is usually sufficiently large to prevent $\rs$ from greatly exceeding $\rIa$.  SNe Ia thus regulates the SMBH accretion rate below the value $\dot{M}_{\rm Ia}$ (eq.~[\ref{eq:eddr_Ia}]), except possibly in high mass elliptical galaxies with large break radii and $\zeta(v_w^{\rm Ia}) \gtrsim \zeta_c$.

\item Heating from SMBH feedback, unlike stellar heating, depends on the SMBH accretion rate, which is itself a function of the heating rate.  Compton heating is generally unimportant compared to stellar wind heating for the average SFH, but it can dominate in the case of an impulsive star burst (Fig.~\ref{fig:vwSourcesImp}).  Although in principle Compton heating is sufficiently high to prevent thermal instabilities when acting alone, i.e. $\dot{M}_{C} \lesssim \dot{M}_{\rm TI}$ (Fig.~\ref{fig:bh_growth}), SMBH feedback may not however be capable of truly stabilizing the flow given its inability to respond instantaneously to local changes in the properties of the gas.  Kinetic feedback could also be important in determining $\dot{M}(M_{\bullet}$) when stellar heating is weak ($\S\ref{sec:kinetic}$) but is more challenging to quantify.
